An Introduction to Rocker:\\ Docker Containers for R

### 2017-11-24 boettiger-eddelbuettel.tex
render and commit

The \texttt{docker\ run} option \texttt{-p} sets the port on which
RStudio\textsuperscript{\textregistered} will appear, for which 8787 is
the default (adding your user to the docker group to avoids the need for
a \texttt{sudo} command to call \texttt{docker}:
\texttt{sudo\ usermod\ -g\ docker\ \$USER}). Many academic and
commercial cloud providers make it possible to execute such code
snippets when a container is launched, without ever needing to
\texttt{ssh} into the machine. The user may log into the server merely
by pasting its IP address or DNS name (followed by the chosen port,
\emph{e.g.}, \texttt{:8787}) into a browser and entering the appropriate
password. This provides the user with a familiar, interactive
environment running on a remote machine while requiring a minimum of
expertise.

@@

which can be useful for administrative tasks such as installing system
dependencies. All Rocker images can also be run as an interactive R,
RScript or bash shell without running RStudio, which can be useful for
batch jobs or for anyone who prefers that environment.

As with any interactive Docker container, users should specify the
terminal (\texttt{-t}) and interactive (\texttt{-i}) flags, (here
combined with interactive as \texttt{-ti}), and specify the desired
executable environment (\emph{e.g.}, \texttt{R}, though other common
options may be \texttt{Rscript} or \texttt{bash}):

@@

Another feature of Rocker containers is the ability to provide a
sandboxed environment, isolated from software and potentially from other
data on the machine. Many users are reluctant to upgrade their suite of
installed packages, which may break their existing code or even their R
environment if the installation goes poorly. However, upgrading packages
and/or the R environment is often necessary to run analysis of a
colleague, or access more recent methods. Rocker offers an easy
solution. For instance, a user can run R code requiring the most recent
versions of R and related packages inside a Rocker container without
having to upgrade their local installations first. Conversely, one could
use Rocker to run code on an older R release with prior versions of R
packages, again without having to make any alteration to one's local R
install. Another common use case is to access a container with support
for particular options such as using gcc or clang compiler sanitizers
\citep{edd_sanitizers}. These require R itself be built with specialized
settings that may not be not available or familiar to many R users on
their native system, but can be easily deployed by pulling the Rocker
images \texttt{rocker/r-devel-san} or
\texttt{rocker/r-devel-ubsan-clang}.

This sandboxing feature is also valuable in the remote computing
context, allowing system administrators to grant users freedom to
install software which requires root privileges inside a container,
while not granting them root access on the host machine. Root access is
required to launch Docker containers, though not to access containers
already running and providing some service such as RStudio. Users
logging into a container through the
RStudio\textsuperscript{\textregistered} interface do not by default
have root privileges, though are able to install R packages. Granting
these users root privileges in the container still leaves them sandboxed
from the host container.\\
Sandboxing also serves an important function in reproducible research by
making it easier to test a specified environment in isolation from the
host machine. Unlike traditional virtual machines, these containers do
not impose a large footprint of reserved resources as a typical host can
easily support 100s of containers \citep{what-docker}.

@@

Having a shared, transparent computational environment created by a
publicly hosted, reproducible recipe facilitates community input into
configuration details. R and many of its packages and related software
can be configured with a wide range of options, compilers, different
linear-algebra libraries and so forth. While this flexibility reflects
varying needs, many users rely on default settings which are most often
are optimized more for simplicity of installation rather than than
performance. The Rocker recipes reflect significant community input on
these choices. This helps create a more finely tuned, optimized
reference implementation of the R environment as well as a platform for
comparing and discussing these concerns which are often overlooked
elsewhere. Issues and Pull Requests on the Rocker repositories on GitHub
attest to some of these discussions and improvements. In particular,
input from the Docker Inc employees through the official approval
process for the \texttt{r-base} image, expertise from the Debian R
maintainer and other Debian developers, and both direct and indirect
feedback from the experience and user-generated documentation from many
early adopters in the R community has helped shape and strengthen the
project over the past few years. Widespread use of the Rocker image
helps promote both testing of these choices and contributions further
tweaking the configuration from many members of the R community.

@@

Users building on the version-tagged images will by default use the MRAN
snapshot mirror \citep{MRAN} associated with the most recent date for
which that image was current. This ensures that a Dockerfile building
\texttt{FROM\ rocker/verse:3.4.1} will only install R package versions
that were available on CRAN on 2017-06-30, that is, the day R 3.4.1 was
released. This default can of course be overwritten in the standard R
manner, \emph{e.g.}, by specifying a different CRAN mirror explicitly in
any command to \texttt{install.packages()}, or by adjusting the default
CRAN mirror in
\texttt{options(repo=\textless{}CRAN-MIRROR\textgreater{})} in an
\texttt{.Rprofile}. Note that the MRAN date associated with the current
release (\emph{e.g.}, \texttt{3.4.2} at the time of writing) will
continue to advance on the Docker-hub image until the next R release.
Software installed from \texttt{apt-get} in these images will come from
the the stable Debian release (\texttt{stretch} or \texttt{jessie}) and
thus not change versions (though it will receive security patches).
Packages installed from BioConductor using the \texttt{bioclite()}
utility will also install the version appropriate to the version of R
found on the system (the Bioconductor semi-annual release model avoids
the need for an MRAN mirror). Users installing packages from GitHub or
other sources can request a specific git release tag or hash for a more
reproducible build, or adopt an alternative approach such as
\texttt{packrat} \citep{packrat}. A more general discussion of the use
and limitations of Docker for computational reproducibility can be found
in \cite{Boettiger2015}.

@@

To make extensions transparent and persistent, Rocker images can be
extended by any user by writing their own Dockerfiles based on an
appropriate Rocker image. The Dockerfiles in the Rocker stack should
themselves provide a simple of example of this, (as described in the
following section).\\
A user begins by selecting an appropriate base image for their needs: if
the RStudio\textsuperscript{\textregistered} interface is desired, a
user might start with \texttt{FROM\ rocker/rstudio}; an image for
testing an R package with compiled code might use
\texttt{FROM\ rocker/r-devel-san}, an image for reproducing a data
analysis will probably select a stable version tag in addition to an
appropriate base library, \emph{e.g.},:
\texttt{FROM\ rocker/tidyverse:3.4.1} Users can easily add additional
software to any running Rocker image using the standard R and Debian
mechanisms. Details on how to extend Rocker images can be found at
\url{https://rocker-project.org}.

@@

The Rocker project consists of a suite of images built automatically by
and hosted on the Docker Hub, \url{https://hub.docker.com/r/rocker}.
Source Dockerfiles, supporting scripts and documentation are hosted on
GitHub under the organization \texttt{rocker-org},
\url{https://github.com/rocker-org}. The issue tracker and pull requests
are used for community input, discussions and contributions to these
images. The Rocker project wiki,
\url{https://github.com/rocker-org/rocker/wiki}, provides a place to
synthesize community-contributed documentation, use-cases, and other
knowledge about using the Rocker images.

@@

The Rocker project aims to provide a small core of Docker images that
serve as convenient `base' images on which other users can build custom
R environments by writing their own Dockerfiles, while also providing a
`batteries included' approach of images that can be used out of the box.
The challenges of balancing diverse needs driven by very different use
cases against the overarching goals of creating images that are still
sufficiently light-weight, easy to use and easy to maintain is a
difficult art. The implementation in both individual Rocker images and
image stacks can never perfect that balance for everyone, but today
reflects the considerable community input and testing over the past few
years.

All Rocker images are based on the Debian Linux distribution. The Debian
platform provides a small base image, the well-known \texttt{apt}
package management system and rich ecosystem of software libraries,
making it the base image of choice for Docker images, including many of
the ``official'' images maintained by Docker's own development team. The
Debian platform is also perhaps the best-supported Linux platform within
the R community, including an active \texttt{r-sig-debian} listserve.
The relatively long period between stable Debian releases (roughly two
years recently) means that software in the Debian stable (\emph{e.g.},
\texttt{debian:jessie}, \texttt{debian:stretch}) releases can lag
significantly behind current releases of popular software, including R.
More recent versions of packages can be found in the pre-release
distribution, \texttt{debian:testing}, while the very latest binary
builds can be found on \texttt{debian:unstable}. The Rocker project can
be largely divided into two stacks which address different needs,
reflected in which Debian distribution they are based on. The first
stack is based on \texttt{debian:testing}. The second, more
recently-introduced stack, is based only on Debian stable releases.
Rocker images always point to specific stable releases (\texttt{jessie},
\texttt{stretch}), and do not use the tag \texttt{debian:stable}, which
is a rolling tag that always points to the most recent stable version.
The different Rocker stacks have different aims and thus provide
different images, as shown in Tables 1 \& 2 below.

@@

The \texttt{debian:testing}-based stack currently includes seven images
actively maintained by the Rocker development team (Table 1).
\texttt{r-base} builds on \texttt{debian:testing}, and the other six in
the stack each build directly from \texttt{r-base}. The \texttt{r-base}
image is unique in that it is designated as the official image for the R
language by the Docker organization itself. This official image is
reviewed and then built by employees of Docker Inc based on a Dockerfile
maintained by the Rocker team. Consequently, users should refer to this
image in Docker commands without an organization namespace, \emph{e.g.},
\texttt{docker\ run\ -ti\ r-base} to access the official image. All
other images in the Rocker project are not individually reviewed and
built by Docker Inc and must be referenced using the \texttt{rocker}
namespace, \emph{e.g.}, \texttt{docker\ run\ -ti\ rocker/r-devel}.

Several of the images in this stack are oriented towards the R
development community: \texttt{r-devel}, \texttt{drd},
\texttt{r-devel-san} \& \texttt{r-devel-ubsan-clang} all add a copy of
the development version of R side-by-side the current release of R
provided by \texttt{r-base}. On these images, the development version is
aliased to \texttt{RD} to distinguish from the current release,
\texttt{R}. As the names suggest, each provide slightly different
configurations. Of particular interest are the images providing
development R built with support for C/C++ address and
undefined-behavior sanitizers, which are somewhat more difficult to
configure \citep{edd_sanitizers}.

@@

\textbf{Build schedule}: The official \texttt{r-base} image is rebuilt
by Docker following any updates to the official \texttt{debian} images
(roughly every few weeks). The rest of the stack uses build triggers
that rebuild the images either whenever \texttt{r-base} is updated, or
the Dockerfile sources are updated on the corresponding GitHub
repository. The only exception in this stack is the \texttt{drd} image,
which is rebuilt each week by a \texttt{cron} trigger.

\begin{longtable}[]{@{}llll@{}}
\caption{The \texttt{debian:testing} image stack}\tabularnewline
\toprule
\begin{minipage}[b]{0.20\columnwidth}\raggedright\strut
image\strut
\end{minipage} & \begin{minipage}[b]{0.50\columnwidth}\raggedright\strut
description\strut
\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\raggedright\strut
size\strut
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
downloads\strut
\end{minipage}\tabularnewline
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{0.20\columnwidth}\raggedright\strut
image\strut
\end{minipage} & \begin{minipage}[b]{0.50\columnwidth}\raggedright\strut
description\strut
\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\raggedright\strut
size\strut
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
downloads\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.20\columnwidth}\raggedright\strut
\href{https://hub.docker.com/r/_/r-base}{r-base}\strut
\end{minipage} & \begin{minipage}[t]{0.50\columnwidth}\raggedright\strut
official image with current version of R\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
254 MB\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
632,000\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.20\columnwidth}\raggedright\strut
\href{https://hub.docker.com/r/rocker/r-devel}{r-devel}\strut
\end{minipage} & \begin{minipage}[t]{0.50\columnwidth}\raggedright\strut
R-devel added side-by-side r-base (using alias \texttt{RD})\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
1 GB\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
4,000\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.20\columnwidth}\raggedright\strut
\href{https://hub.docker.com/r/rocker/drd}{drd}\strut
\end{minipage} & \begin{minipage}[t]{0.50\columnwidth}\raggedright\strut
lightweight r-devel, built weekly\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
571 MB\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
4,000\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.20\columnwidth}\raggedright\strut
\href{https://hub.docker.com/r/rocker/r-devel-san}{r-devel-san}\strut
\end{minipage} & \begin{minipage}[t]{0.50\columnwidth}\raggedright\strut
as r-devel, but built with compiler sanitizers\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
1.1 GB\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
1,000\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.20\columnwidth}\raggedright\strut
\href{https://hub.docker.com/r/rocker/r-devel-ubsan-clang}{r-devel-ubsan-clang}\strut
\end{minipage} & \begin{minipage}[t]{0.50\columnwidth}\raggedright\strut
Sanitizers, clang c compiler (instead of gcc)\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
1.1 GB\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
525\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.20\columnwidth}\raggedright\strut
\href{https://hub.docker.com/r/rocker/r-devel-san}{rstudio:testing}\strut
\end{minipage} & \begin{minipage}[t]{0.50\columnwidth}\raggedright\strut
rstudio on debian:testing\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
1.1 GB\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
1,000\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.20\columnwidth}\raggedright\strut
\href{https://hub.docker.com/r/rocker/shiny}{shiny}\strut
\end{minipage} & \begin{minipage}[t]{0.50\columnwidth}\raggedright\strut
shiny-server on r-base\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
409 MB\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
123,000\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

\begin{longtable}[]{@{}llll@{}}
\caption{The \texttt{rocker-versioned} stack of images}\tabularnewline
\toprule
image & description & size & downloads\tabularnewline
\midrule
\endfirsthead
\toprule
image & description & size & downloads\tabularnewline
\midrule
\endhead
\href{https://hub.docker.com/r/rocker/r-ver}{r-ver} & Version-stable
base R \& src build tools & 219 MB & 6,000\tabularnewline
\href{https://hub.docker.com/r/rocker/rstudio}{rstudio} & Adds rstudio &
334 MB & 314,000\tabularnewline
\href{https://hub.docker.com/r/rocker/tidyverse}{tidyverse} & Adds
tidyverse \& devtools & 656 MB & 83,000 \footnote{This figure includes
  49,000 downloads under the earlier name \texttt{hadleyverse}.}\tabularnewline
\href{https://hub.docker.com/r/rocker/verse}{verse} & Adds java, tex \&
publishing-related packages & 947 MB & 9,000\tabularnewline
\href{https://hub.docker.com/r/rocker/geospatial}{geospatial} & Adds
geospatial libraries & 1.3 GB & 4,000\tabularnewline
\bottomrule
\end{longtable}

@@

This stack emphasizes stability and reproducibility of the Docker build.
This stack was introduced much more recently (November 2016) in response
to considerable user input and requests. The key feature of this stack
is the ability to run older versions of R along with the
then-contemporaneous versions of R packages. A user specifies the
version desired using an image tag, \emph{e.g.},
\texttt{rocker/r-ver:3.3.1} will refer to an image with R version 3.3.1
installed. Omitting the tag is equivalent to using the tag
\texttt{latest}, which, as the name implies, will always point to an
image using the current R release. Thus, users wanting to create
downstream Dockerfiles which are based on the current release at time
(but will continue to reconstruct the same environment in the future
after newer R versions are released) should explicitly include the
corresponding version tag, \emph{e.g.}, \texttt{rocker/r-ver:3.4.2} at
the time of writing, and not the \texttt{latest} tag. Users can also run
the current development version of R using the tag \texttt{devel}, which
is built nightly from R-devel sources from \texttt{subversion}.

@@

\textbf{Build schedule}: All images are built automatically from their
corresponding Dockerfiles (found in the GitHub repositories
\texttt{rocker-org/rocker-versioned} and
\texttt{rocker-org/geospatial}). A \texttt{cron} job sends nightly build
triggers to Docker Hub to rebuild the \texttt{latest} and \texttt{devel}
tagged images throughout the stack. To decrease load on the hub, build
triggers for the numeric version tags are sent monthly. Although the
Dockerfiles for older R versions installs an almost-identical software
environment every time, the monthly rebuilding of these images on Docker
Hub ensures they continue to receive Debian security updates from
upstream, and proves the build recipe still executes successfully. Note
that rebuilding images with software from external repositories never
produces a bit-wise identical image, and thus the image identifier hash
will change at each build.

>>>>

The \texttt{docker\ run} option \texttt{-p} sets the port on which
RStudio\textsuperscript{\textregistered} will appear, for which 8787 is
the default (adding your user to the docker group avoids the need for a
\texttt{sudo} command to call \texttt{docker}:
\texttt{sudo\ usermod\ -g\ docker\ \$USER}). Many academic and
commercial cloud providers make it possible to execute such code
snippets when a container is launched, without ever needing to
\texttt{ssh} into the machine. The user may log into the server merely
by pasting its IP address or DNS name (followed by the chosen port,
\emph{e.g.}, \texttt{:8787}) into a browser and entering the appropriate
password. This provides the user with a familiar, interactive
environment running on a remote machine while requiring a minimum of
expertise.

@@

\noindent which can be useful for administrative tasks such as
installing system dependencies. All Rocker images can also be run as an
interactive R, RScript or bash shell without running RStudio, which can
be useful for batch jobs or for anyone who prefers that environment.

As with any interactive Docker container, users should specify the
terminal (\texttt{-t}) and interactive (\texttt{-i}) flags, (here
combined with interactive as \texttt{-ti}), and specify the desired
executable environment (\emph{e.g.}, \texttt{R}, though other common
options could be \texttt{Rscript} or \texttt{bash}):

@@

Another feature of Rocker containers is the ability to provide a
sandboxed environment, isolated from software and potentially from other
data on the machine. Many users are reluctant to upgrade their suite of
installed packages, which may break their existing code or even their R
environment if the installation goes poorly. However, upgrading packages
and/or the R environment is often necessary to run analyses from a
colleague, or access more recent methods. Rocker offers an easy
solution. For instance, a user can run R code requiring the most recent
versions of R and related packages inside a Rocker container without
having to upgrade their local installations first. Conversely, one could
use Rocker to run code on an older R release with prior versions of R
packages, again without having to make any alteration to one's local R
install. Another common use case is to access a container with support
for particular options such as using gcc or clang compiler sanitizers
\citep{edd_sanitizers}. These require R itself be built with specialized
settings that may not be not available or familiar to many R users on
their native system, but can be easily deployed by pulling the Rocker
images \texttt{rocker/r-devel-san} or
\texttt{rocker/r-devel-ubsan-clang}.

This sandboxing feature is also valuable in the remote computing
context, allowing system administrators to grant users freedom to
install software which requires root privileges inside a container,
while not granting them root access on the host machine. Root access is
required to launch Docker containers, though not to access containers
already running and providing some service such as RStudio. Users
logging into a container through the
RStudio\textsuperscript{\textregistered} interface do not by default
have root privileges, though are able to install R packages. Granting
these users root privileges in the container still leaves them sandboxed
from the host container.\\
Sandboxing also serves an important function in reproducible research by
making it easier to test a specified environment in isolation from the
host machine. Unlike traditional virtual machines, containers do not
impose a large footprint of reserved resources as a typical host can
easily support 100s of containers \citep{what-docker}.

@@

Having a shared, transparent computational environment created by a
publicly hosted, reproducible recipe facilitates community input into
configuration details. R and many of its packages and related software
can be configured with a wide range of options, compilers, different
linear-algebra libraries and so forth. While this flexibility reflects
varying needs, many users rely on default settings which are most often
are optimized more for simplicity of installation rather than than
performance. The Rocker recipes reflect significant community input on
these choices. This helps create a more finely tuned, optimized
reference implementation of the R environment as well as a platform for
comparing and discussing these concerns which are often overlooked
elsewhere. Issues and Pull Requests on the Rocker repositories on GitHub
attest to some of these discussions and improvements. In particular,
input from the Docker Inc employees through the official approval
process for the \texttt{r-base} image, expertise from the Debian R
maintainer and other Debian developers, and both direct and indirect
feedback from the experience and user-generated documentation from many
early adopters in the R community has helped shape and strengthen the
project over the past few years. Widespread use of the Rocker image
helps promote both testing of these choices and contributions, further
tweaking the configuration from many members of the R community.

@@

Users building on the version-tagged images will by default use the MRAN
snapshot mirror \citep{MRAN} associated with the most recent date for
which that image was current. This ensures that a Dockerfile building
\texttt{FROM\ rocker/verse:3.4.1} will only install R package versions
that were available on CRAN on 2017-06-30, \emph{i.e.}, the day R 3.4.1
was released. This default can of course be overwritten in the standard
R manner, \emph{e.g.}, by specifying a different CRAN mirror explicitly
in any command to install packages, \emph{e.g.},
\texttt{install.packages()}, or by adjusting the default CRAN mirror in
\texttt{options(repo=\textless{}CRAN-MIRROR\textgreater{})} in an
\texttt{.Rprofile}. Note that the MRAN date associated with the current
release (\emph{e.g.}, \texttt{3.4.2} at the time of writing) will
continue to advance on the Docker-hub image until the next R release.
Software installed from \texttt{apt-get} in these images will come from
the the stable Debian release (\texttt{stretch} or \texttt{jessie}) and
thus not change versions (though it will receive security patches).
Packages installed from BioConductor using the \texttt{bioclite()}
utility will also install the version appropriate to the version of R
found on the system (the Bioconductor semi-annual release model avoids
the need for an MRAN mirror). Users installing packages from GitHub or
other sources can request a specific git release tag or hash for a more
reproducible build, or adopt an alternative approach such as
\texttt{packrat} \citep{packrat}. A more general discussion of the use
and limitations of Docker for computational reproducibility can be found
in \cite{Boettiger2015}.

@@

To make extensions transparent and persistent, Rocker images can be
extended by any user by writing their own Dockerfiles based on an
appropriate Rocker image. The Dockerfiles in the Rocker stack should
themselves provide a simple example of this, (as described in the
following section).\\
A user begins by selecting an appropriate base image for their needs: if
the RStudio\textsuperscript{\textregistered} interface is desired, a
user might start with \texttt{FROM\ rocker/rstudio}; an image for
testing an R package with compiled code might use
\texttt{FROM\ rocker/r-devel-san}, and an image for reproducing a data
analysis will probably select a stable version tag in addition to an
appropriate base library, \emph{e.g.},:
\texttt{FROM\ rocker/tidyverse:3.4.1} Users can easily add additional
software to any running Rocker image using the standard R and Debian
mechanisms. Details on how to extend Rocker images can be found at
\url{https://rocker-project.org}.

@@

The Rocker project consists of a suite of images built automatically by
and hosted on the Docker Hub, \url{https://hub.docker.com/r/rocker}.
Source Dockerfiles, supporting scripts and documentation are hosted on
GitHub under the organization \texttt{rocker-org},
\url{https://github.com/rocker-org}. The issue tracker and pull requests
are used for community input, discussions, and contributions to these
images. The Rocker project wiki,
\url{https://github.com/rocker-org/rocker/wiki}, provides a place to
synthesize community-contributed documentation, use-cases, and other
knowledge about using the Rocker images.

@@

The Rocker project aims to provide a small core of Docker images that
serve as convenient `base' images on which other users can build custom
R environments by writing their own Dockerfiles, while also providing a
`batteries included' approach to images that can be used out of the box.
The challenges of balancing diverse needs driven by very different use
cases against the overarching goals of creating images that are still
sufficiently light-weight, easy to use, and easy to maintain is a
difficult art. The implementation in both individual Rocker images and
image stacks can never perfect that balance for everyone, but today
reflects the considerable community input and testing over the past few
years.

All Rocker images are based on the Debian Linux distribution. The Debian
platform provides a small base image, the well-known \texttt{apt}
package management system, and a rich ecosystem of software libraries,
making it the base image of choice for Docker images, including many of
the ``official'' images maintained by Docker's own development team. The
Debian platform is also perhaps the best-supported Linux platform within
the R community, including an active \texttt{r-sig-debian} listserve.
The relatively long period between stable Debian releases (roughly two
years recently) means that software in the Debian stable (\emph{e.g.},
\texttt{debian:jessie}, \texttt{debian:stretch}) releases can lag
significantly behind current releases of popular software, including R.
More recent versions of packages can be found in the pre-release
distribution, \texttt{debian:testing}, while the very latest binary
builds can be found on \texttt{debian:unstable}. The Rocker project can
be largely divided into two stacks which address different needs,
reflected in which Debian distribution they are based on. The first
stack is based on \texttt{debian:testing}. The second, more
recently-introduced stack, is based only on Debian stable releases.
Rocker images always point to specific stable releases (\texttt{jessie},
\texttt{stretch}), and do not use the tag \texttt{debian:stable}, which
is a rolling tag that always points to the most recent stable version.
The different Rocker stacks have different aims and thus provide
different images, as shown in Tables 1 \& 2 below.

@@

The \texttt{debian:testing}-based stack currently includes seven images
actively maintained by the Rocker development team (Table 1).
\texttt{r-base} builds on \texttt{debian:testing}, and the other six in
the stack each build directly from \texttt{r-base}. The \texttt{r-base}
image is unique in that it is designated as the official image for the R
language by the Docker organization itself. This official image is
reviewed and then built by employees of Docker Inc. based on a
Dockerfile maintained by the Rocker team. Consequently, users should
refer to this image in Docker commands without an organization
namespace, \emph{e.g.}, \texttt{docker\ run\ -ti\ r-base} to access the
official image. All other images in the Rocker project are not
individually reviewed and built by Docker Inc. and must be referenced
using the \texttt{rocker} namespace, \emph{e.g.},
\texttt{docker\ run\ -ti\ rocker/r-devel}.

Several of the images in this stack are oriented towards the R
development community: \texttt{r-devel}, \texttt{drd},
\texttt{r-devel-san}, and \texttt{r-devel-ubsan-clang} which all add a
copy of the development version of R side-by-side to the current release
of R provided by \texttt{r-base}. On these images, the development
version is aliased to \texttt{RD} to distinguish from the current
release, \texttt{R}. As the names suggest, each provide slightly
different configurations. Of particular interest are the images
providing development R built with support for C/C++ address and
undefined-behavior sanitizers, which are somewhat difficult to configure
\citep{edd_sanitizers}.

@@

\textbf{Build schedule}: The official \texttt{r-base} image is rebuilt
by Docker following any updates to the official \texttt{debian} images
(roughly every few weeks). The rest of the stack uses build triggers
that rebuild the images whenever \texttt{r-base} is updated or the
Dockerfile sources are updated on the corresponding GitHub repository.
The only exception in this stack is the \texttt{drd} image, which is
rebuilt each week by a \texttt{cron} trigger.

\begin{longtable}[]{@{}llll@{}}
\caption{The \texttt{debian:testing} image stack}\tabularnewline
\toprule
\begin{minipage}[b]{0.20\columnwidth}\raggedright\strut
image\strut
\end{minipage} & \begin{minipage}[b]{0.50\columnwidth}\raggedright\strut
description\strut
\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\raggedright\strut
size\strut
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
downloads\strut
\end{minipage}\tabularnewline
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{0.20\columnwidth}\raggedright\strut
image\strut
\end{minipage} & \begin{minipage}[b]{0.50\columnwidth}\raggedright\strut
description\strut
\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\raggedright\strut
size\strut
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
downloads\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.20\columnwidth}\raggedright\strut
\href{https://hub.docker.com/r/_/r-base}{r-base}\strut
\end{minipage} & \begin{minipage}[t]{0.50\columnwidth}\raggedright\strut
official image with current version of R\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
254 MB\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
632,000\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.20\columnwidth}\raggedright\strut
\href{https://hub.docker.com/r/rocker/r-devel}{r-devel}\strut
\end{minipage} & \begin{minipage}[t]{0.50\columnwidth}\raggedright\strut
R-devel added side-by-side to r-base (using alias \texttt{RD})\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
1 GB\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
4,000\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.20\columnwidth}\raggedright\strut
\href{https://hub.docker.com/r/rocker/drd}{drd}\strut
\end{minipage} & \begin{minipage}[t]{0.50\columnwidth}\raggedright\strut
lightweight r-devel, built weekly\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
571 MB\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
4,000\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.20\columnwidth}\raggedright\strut
\href{https://hub.docker.com/r/rocker/r-devel-san}{r-devel-san}\strut
\end{minipage} & \begin{minipage}[t]{0.50\columnwidth}\raggedright\strut
as r-devel, but built with compiler sanitizers\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
1.1 GB\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
1,000\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.20\columnwidth}\raggedright\strut
\href{https://hub.docker.com/r/rocker/r-devel-ubsan-clang}{r-devel-ubsan-clang}\strut
\end{minipage} & \begin{minipage}[t]{0.50\columnwidth}\raggedright\strut
sanitizers, clang c compiler (instead of gcc)\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
1.1 GB\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
525\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.20\columnwidth}\raggedright\strut
\href{https://hub.docker.com/r/rocker/r-devel-san}{rstudio:testing}\strut
\end{minipage} & \begin{minipage}[t]{0.50\columnwidth}\raggedright\strut
rstudio on debian:testing\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
1.1 GB\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
1,000\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.20\columnwidth}\raggedright\strut
\href{https://hub.docker.com/r/rocker/shiny}{shiny}\strut
\end{minipage} & \begin{minipage}[t]{0.50\columnwidth}\raggedright\strut
shiny-server on r-base\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
409 MB\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
123,000\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

\begin{longtable}[]{@{}llll@{}}
\caption{The \texttt{rocker-versioned} stack of images}\tabularnewline
\toprule
image & description & size & downloads\tabularnewline
\midrule
\endfirsthead
\toprule
image & description & size & downloads\tabularnewline
\midrule
\endhead
\href{https://hub.docker.com/r/rocker/r-ver}{r-ver} & version-stable
base R \& src build tools & 219 MB & 6,000\tabularnewline
\href{https://hub.docker.com/r/rocker/rstudio}{rstudio} & adds rstudio &
334 MB & 314,000\tabularnewline
\href{https://hub.docker.com/r/rocker/tidyverse}{tidyverse} & adds
tidyverse \& devtools & 656 MB & 83,000 \footnote{This figure includes
  49,000 downloads under the earlier name \texttt{hadleyverse}.}\tabularnewline
\href{https://hub.docker.com/r/rocker/verse}{verse} & adds java, tex \&
publishing-related packages & 947 MB & 9,000\tabularnewline
\href{https://hub.docker.com/r/rocker/geospatial}{geospatial} & adds
geospatial libraries & 1.3 GB & 4,000\tabularnewline
\bottomrule
\end{longtable}

@@

This stack emphasizes stability and reproducibility of the Docker build.
This stack was introduced much more recently (November 2016) in response
to considerable user input and requests. The key feature of this stack
is the ability to run older versions of R along with the
then-contemporaneous versions of R packages. A user specifies the
version desired using an image tag, \emph{e.g.},
\texttt{rocker/r-ver:3.3.1} will refer to an image with R version 3.3.1
installed. Omitting the tag is equivalent to using the tag
\texttt{latest}, which, as the name implies, will always point to an
image using the current R release. Thus, users who want to create
downstream Dockerfiles, which are based on the current release at the
time (but will continue to reconstruct the same environment in the
future after newer R versions are released), should explicitly include
the corresponding version tag, \emph{e.g.}, \texttt{rocker/r-ver:3.4.2}
at the time of writing, and not the \texttt{latest} tag. Users can also
run the current development version of R using the tag \texttt{devel},
which is built nightly from R-devel sources from \texttt{subversion}.

@@

\textbf{Build schedule}: All images are built automatically from their
corresponding Dockerfiles (found in the GitHub repositories
\texttt{rocker-org/rocker-versioned} and
\texttt{rocker-org/geospatial}). A \texttt{cron} job sends nightly build
triggers to Docker Hub to rebuild the \texttt{latest} and \texttt{devel}
tagged images throughout the stack. To decrease load on the hub, build
triggers for the numeric version tags are sent monthly. Although the
Dockerfiles for older R versions install an almost-identical software
environment every time, the monthly rebuilding of these images on Docker
Hub ensures they continue to receive Debian security updates from
upstream, and proves the build recipe still executes successfully. Note
that rebuilding images with software from external repositories never
produces a bit-wise identical image, and thus the image identifier hash
will change at each build.
